---
title: "Data Import"
author: "Samantha Toet"
date: "7/23/2020"
output: html_document
---

```{r setup, echo=FALSE}
library(tidyverse)
```

## Intro

From [11. Data Import](https://r4ds.had.co.nz/data-import.html)

The `readr` package turns flat files into dataframes. 

The output is a tibble, they don't convert character vectors to factors, or use row names. 

### Practice

**1. What function would you use to read a file where fields were separated with | ?"**

`readr::read_delim("filename", delim = "|")`

**2. Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?**

- `col_names` defines column names (bool or char vector supplied)
- `col_types` defines types of columns 
- `locale` controls location-based defaults (ex. currency)
- `na` defines values to treat as NA
- `quoted_na` defines missing values in strings
- `quote` single character used to quote strings
- `trim_ws` trims leading and trailing white space
- `n_max` max number of records to read in
- `guess_max` max number of records to use for guessing column types
- `progress` display progress bar
- `skip_empty_rows` how to handle empty rows

**3. What are the most important arguments to `read_fwf()`?**

The most important arguments are `col_positions` (if created by `fwf_empty()`, `fwf_widths()` or `fwf_positions()`). To read in only selected fields, use `fwf_positions()`. If the width of the last column is variable (a ragged fwf file), supply the last end position as NA.

## Parsing a vector

Parse functions take a character vector and return either a logical, integer, or date. 

Each type has its own set of complications due to encodings and locales. 

### Practice

**1. What are the most important arguments to `locale()`?**

The most important arguments to `locale()` are `date_format()`, which defines how and in what language dates appear, and `decimal_mark` which determines how vales are grouped. 

**2. What happens if you try and set `decimal_mark` and `grouping_mark` to the same character? What happens to the default value of `grouping_mark` when you set `decimal_mark` to “,”? What happens to the default value of `decimal_mark` when you set the `grouping_mark` to “.”?**

```{r error=TRUE}

# Different decimal and grouping parses as expected:
parse_double(c("1.23", "1234.5"), locale = locale(decimal_mark = ".", grouping_mark = ","))

# Same decimal and grouping throws and error:
parse_double(c("1.23", "1234.5"), locale = locale(decimal_mark = ".", grouping_mark = "."))

# If decimal is set to , then grouping becomes . and vice-versa
parse_double(c("1.23", "1234.5"), locale = locale(decimal_mark = ","))
parse_double(c("1.23", "1234.5"), locale = locale(grouping_mark = "."))
```

**3. I didn’t discuss the `date_format` and `time_format` options to `locale()`. What do they do? Construct an example that shows when they might be useful.**

The `date_format` and `time_format` arguments determine the default date and time formats. This is helpful if you're working with data that's not US encoded. For example:

```{r}
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```

**4. If you live outside the US, create a new locale object that encapsulates the settings for the types of file you read most commonly.**

Let's pretend I live in Montreal, Canada:

```{r}
locale("fr", tz = "EST")
```

**5. What’s the difference between `read_csv()` and `read_csv2()`?**

They both read in CSV files, however `read_csv2()` uses ; for the field separator and , for the decimal point. This is common formatting in some European countries. 

**6. What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.**

Western Europe encodings (ISO 8859-1, ISO 8859-15 and cp1252)

Chinese encodings (GB2312, …), and then come Korean (EUC-KR), Cyrillic (cp1251, KOI8-R, …), East Europe (cp1250, ISO-8859-2), Arabic (cp1256, ISO-8859-6), etc.

**7. Generate the correct format string to parse each of the following dates and times:**

```{r}
d1 <- "January 1, 2010"
parse_date(d1, "%B %m, %Y")

d2 <- "2015-Mar-07"
parse_date(d2, "%Y-%b-%e")

d3 <- "06-Jun-2017"
parse_date(d3, "%e-%b-%Y")

d4 <- c("August 19 (2015)", "July 1 (2015)")
parse_date(d4, "%B %d (%Y)")

d5 <- "12/30/14" # Dec 30, 2014
parse_date(d5, "%m/%d/%y")

t1 <- "1705"
parse_time(t1, "%H%M")

t2 <- "11:15:10.12 PM"
parse_time(t2, "%I:%M:%OS %p")
```

## Parsing & writing files

Essentially, `readr` guesses the type of each column by using the first 1000 rows as a heuristic. 

When it comes to writing files, `write_csv` encodes strings in UTF-8 and dates in ISO8601. 

Class information is lost when writing to a CSV, so other options are `write_rds` or `write_feather`
