---
title: "Relational Data"
author: "Samantha Toet"
date: "10/14/2020"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

## Intro

Two types of joins:

- **Mutating** joins, which add new variables to one data frame from matching observations in another.

- **Filtering** joins, which filter observations from one data frame based on whether or not they match an observation in the other table.

Set operations treat observations as if they were set elements.

### Practice 

**1. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?**

We'd want to track each unique plane's origin and destination locations based on the actual departure and arrival times. The tables that would need to be joined for this are `airports` for timezones, `flights` for locations and times, and `planes` for each plane's unique `tailnum`. 

**2. I forgot to draw the relationship between `weather` and `airports`. What is the relationship and how should it appear in the diagram?**

`weather` connects to `airports` on the `origin` and `faa` variables. 

**3. `weather` only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with `flights`?**

It would also define the `dest` relation. 

**4. We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?**

It would be a dataframe that would have two variables to represent date, `month` and `day`, and then another variable with more detail about why that date is considered special. This dataframe would then be able to join with flights and weather on the `month` and `day` values. 

## Keys

There are two types of keys:

- A **primary* key uniquely identifies an observation in its own table. For example, planes$tailnum is a primary key because it uniquely identifies each plane in the planes table.

- A **foreign** key uniquely identifies an observation in another table. For example, flights$tailnum is a foreign key because it appears in the flights table where it matches each flight to a unique plane.

If a table lacks a primary key, it’s sometimes useful to add one with `mutate()` and `row_number()`. That makes it easier to match observations if you’ve done some filtering and want to check back in with the original data. This is called a **surrogate** key.

### Practice

**1. Add a surrogate key to `flights`.**

```{r}
flights %<>%
  mutate(id = row_number())
```

**2. Identify the keys in the following datasets:** 

- `Lahman::Batting`: playerID, yearID, and teamID
- `babynames::babynames`: year, name
- `nasaweather::atmos`: lat, lon, year, month
- `fueleconomy::vehicles`: id, make, model, year
- `ggplot2::diamonds`: carat, cut, clarity

## Mutating Joins

Matches observations by their keys, then copies across variables from one table to the other.

An **inner join** keeps observations that appear in both tables. 

An **outer join* keeps observations that appear in at least one of the tables. There are three types of outer joins:

- A left join keeps all observations in x.
- A right join keeps all observations in y.
- A full join keeps all observations in x and y.

### Practice

**1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays.**

```{r}
flights %>%
  group_by(dest) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = T)) %>%
  left_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(size = avg_delay, alpha = avg_delay)) +
    coord_quickmap() 
```

**2. Add the location of the origin and destination (i.e. the `lat` and `lon`) to `flights`.**

```{r}
flights %>%
  left_join(airports, by = c("origin" = "faa")) %>%
  select(-alt, -tz, -dst, -tzone) %>%
  rename(origin_lat = lat, origin_lon = lon) %>%
  left_join(airports, by = c("dest" = "faa")) %>%
  select(-alt, -tz, -dst, -tzone) %>%
  rename(dest_lat = lat, dest_lon = lon)
```

**3. Is there a relationship between the age of a plane and its delays?**

```{r}
flights %>%
  group_by(tailnum) %>%
  summarise(avg_arr_delay = mean(arr_delay)) %>%
  left_join(planes, by = "tailnum") %>%
  mutate(age = 2020 - year) %>%
  ggplot(aes(age, avg_arr_delay, fill = manufacturer)) +
  geom_col()
```

**4. What weather conditions make it more likely to see a delay?**

```{r}
flights %>%
  group_by(tailnum) %>%
  mutate(avg_arr_delay = mean(arr_delay)) %>%
  inner_join(weather, by = c("origin" = "origin", "year" = "year", "month" = "month", "day" = "day", "hour" = "hour")) %>%
  ggplot(aes(visib, avg_arr_delay)) +
  geom_point(aes(alpha = precip))
```

## Filtering Joins

There are two types of filtering joins:

- `semi_join(x, y)` keeps all observations in x that have a match in y.
- `anti_join(x, y)` drops all observations in x that have a match in y.

### Practice

**1. What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)**

```{r}
flights %>%
  filter(is.na(tailnum) & is.na(air_time))
```

The flights that have a missing `tailnum` also have missing departure and arrival times. This means that the flights never actually occured. 

**2. Filter flights to only show flights with planes that have flown at least 100 flights.**

```{r}
flights100 <- flights %>%
  filter(!is.na(tailnum)) %>%
  group_by(tailnum) %>%
  count() %>%
  filter(n >= 100)

flights %>%
  semi_join(flights100, by = "tailnum")
```

**3. Combine `fueleconomy::vehicles` and `fueleconomy::common` to find only the records for the most common models.**

```{r}
vehicles <- fueleconomy::vehicles
common <- fueleconomy::common

vehicles %>%
  semi_join(common, by = c("make" = "make", "model" = "model"))
```

**5. What does `anti_join(flights, airports, by = c("dest" = "faa"))` tell you? What does `anti_join(airports, flights, by = c("faa" = "dest"))` tell you?**

When written using the pipe operator, the meaning appears clearer: 

```{r}
flights %>%
  anti_join(airports, by = c("dest" = "faa"))
```

The above returns all rows from flights without a match in airports. 

```{r}
airports %>%
  anti_join(flights, by = c("faa" = "dest"))
```

The above returns all rows from airports without a match in flights. 

**6. You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.**
