---
title: "21 - iterate"
author: "Samantha Toet"
date: "10/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Intro

Iteration help when you need to do the same thing to multiple inputs, i.e. repeating the same operation on different columns, or on different datasets. 

Functional programming (FP) offers tools to extract out this duplicated code, so each common for loop pattern gets its own function.

---- skipping for loop section ----

## map()

### Practice

1. Write code that uses one of the map functions to:

a. Compute the mean of every column in `mtcars`.

```{r}
mtcars %>%
  map(., mean)
```

b. Determine the type of each column in `nycflights13::flights`.

```{r}
nycflights13::flights %>%
  map(., class)
```

c. Compute the number of unique values in each column of `iris`.

```{r}
iris %>%
  map(., ~ length(unique(.)))
```

d. Generate 10 random normals from distributions with means of -10, 0, 10, and 100.

```{r}
x <- list(10, 0, 10, 100) %>%
  map(. ~ rnrom)

```

2. 
