---
title: "Workflows: Basics"
author: "Samantha Toet"
date: "7/5/2018"
output: html_document
---

## Coding Basics

From [4. Workflows: Basic](http://r4ds.had.co.nz/workflow-basics.html)

*Ch. 2 in the printed copy*

- assignments: dont use =
- naming conventions: snake case
- calling functions: If you want more help, press `F1` to get all the details in the help tab in the lower right pane.
- the continuiation character `+`: tells you that R is waiting for more input; it doesn’t think you’re done yet. Usually that means you’ve forgotten either a `"` or a `)`. Either add the missing pair, or press `ESC` to abort the expression and try again.
- print to screen by surrounding in parantheses 

### Practice:

**1. Why does this code not work?**
```{r example, include=TRUE, echo=TRUE, error=TRUE}
my_variable <- 10
my_varıable
```

**Answer**: R is very picky about spelling and capitalization. The two variables must be spelled and formatted identially. In the above example, the i's in *variable* are different. 

**2. Tweak each of the following R commands so that they run correctly:**

```{r, include=TRUE, echo=TRUE, error=TRUE, warning=F, message=FALSE}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

fliter(mpg, cyl = 8)
filter(diamond, carat > 3)
```

**Answer**: see below output
```{r, include=TRUE, echo=TRUE, error=TRUE, warning=F, message=FALSE}
library(tidyverse)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + # slightly cleaner
  geom_point()

filter(mpg, cyl == 8) # filter spelled correctly and correct boolean token used

ggplot2::diamonds # dataset loaded 
filter(diamonds, carat > 3) # diamonds dataset correctly spelled 
```

**3. Press `Alt` + `Shift` + `K`. What happens? How can you get to the same place using the menus?**

**Answer**: 
Pressing `Alt` + `Shift` + `K` bring up the Keyboard Shortcut Quick Reference. To get to the same place using the menus you'd have to select Tools, then Keyboard Shortcuts Help.  
