---
title: "Tidy"
author: "Samantha Toet"
date: "9/14/2020"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
```

## Intro

From [12. Tidy data](https://r4ds.had.co.nz/tidy-data.html)

Three rules of tidy data:

1.) Each variable must have it's own column
2.) Each observation must have it's own row
3.) Each value must have it's own cell

### Practice

**1. Using prose, describe how the variables and observations are organised in each of the sample tables.**

```{r}
table1
```

table1 can be decribed as tidy. 

```{r}
table2
```

Although each variable has a column, table2 is not tidy because each observations is not recorded in a single row (country/year x2). 

```{r}
table3
```

table3 is not tidy because there are multiple vales per cell in the rate variable. It also has a single observation (country/year) spread across rows. 

```{r}
# Spread across two tibbles
table4a # cases
table4b  # population
```

tables4a and 4b is not tidy because they are spread across two tibbles.

## Pivoting

Pivot LONGER when column names are not names of variables, but values of a variable. For example: 

```{r, error=FALSE}
# pivot_longer(c("cols with names that are values not variables"),
#   names_to = "new column name that the above col names will be values in",
#   values_to = "new column name that the values in the first cols will become values in")
```
  
Pivot WIDER when an observation is scattered across multiple rows. For example: 

```{r, error=FALSE}
# pivot_wider(names_from = "col that has diff variables as values",
#   values_from = "values from first col")
```

### Practice

**1. Why are `pivot_longer()` and `pivot_wider()` not perfectly symmetrical? Carefully consider the following example:** 

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016), # class = dbl
  half  = c(   1,    2,     1,    2), # class = dbl
  return = c(1.88, 0.59, 0.92, 0.17) # class = dbl
)

stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% # class stays dbl for all 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return") # class for year is character
```

**(Hint: look at the variable types and think about column names.) `pivot_longer()` has a `names_ptype` argument, e.g. `names_ptype = list(year = double())`. What does it do?**

When the tibble is pivoted longer, the class for the year variable is changed from a dbl to a character. The `names_ptype` argument preserves the class. 

**2. Why does this code fail?**

```{r error=TRUE}
table4a %>% 
  pivot_longer(c(1999, 2000), names_to = "year", values_to = "cases")
```

The above code fails because `pivot_longer` expects either character column names or column locations by index number. Since 1999 and 2000 are written as numeric, `pivot_longer` thinks those are the column numbers, as in it is looking for the 1999th and 2000th column, and since those don't exist, it throws an error. 

To resolve this issue, put the numeric column names as characters:

```{r}
table4a %>% 
  pivot_longer(c("1999", "2000"), names_to = "year", values_to = "cases")
```

**3. What would happen if you widen this table? Why? How could you add a new column to uniquely identify each value?**

```{r}
people <- tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

Because `people` has a column named values, `pivot_wider` throws an error:
```{r}
people %>%
  pivot_wider(names_from = c("names"), values_from = c("values"))

```

To resolve this, you need to rename a column:

```{r}
people %>% 
  mutate(measurements = values) %>%
  pivot_wider(names_from = c("names"), values_from = c("values"))
```

**4. Tidy the simple tibble below. Do you need to make it wider or longer? What are the variables?**

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

We need to make the table longer so that each row is a unique count. The variables should then be `count`, `gender`, `pregnant`.

```{r}
preg %>%
  pivot_longer(c(male, female), 
               names_to = "gender", 
               values_to = "count")

```

## Separating and uniting

**1. What do the `extra` and `fill` arguments do in `separate()`? Experiment with the various options for the following two toy datasets.**

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))
```

In the above example, the value "g" gets dropped when we separate into 3 columns. To resolve this, we would want to use the `extra` argument to determine what to do with the extra value. We can either emit a warning, drop the extra values (the default), or merge them. For example:

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")
```

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```

In the above example, there are not enough values in the second row to separate into three columns, so one of the values is `NA`. To resolve this, we would want to use the `fill` argument to determine what do to when there are not enough values. We can either emit a warning, fill with missing values on the right (the default), or fill with missing values on the left. For example:

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")
```

**2. Both `unite()` and `separate()` have a `remove` argument. What does it do? Why would you set it to `FALSE`?**

The `remove` argument lets you remove the input column from the output dataframe. To create a new variable, but keep the old one, you would set it to false. For example:

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), remove = FALSE)
```

**3. Compare and contrast `separate()` and `extract()`. Why are there three variations of separation (by position, by separator, and with groups), but only one unite?**

The `extract()` function lets you use regular expressions to define different groups and create new columns, where `separate()` is more limited. With `extract()` and `separate()` only one column can be chosen, but there are many choices how to split that single column into different columns. With `unite()`, there are many choices as to which columns to include, but only onechoice as to how to combine their contents into a single vector. 

## Missing values

There are two types of missing values:

- Explicitly, i.e. flagged with NA.
- Implicitly, i.e. simply not present in the data.

An explicit missing value is the presence of an absence; an implicit missing value is the absence of a presence.

### Practice

**1. Compare and contrast the `fill` arguments to `pivot_wider()` and `complete()`.**

In `pivot_wider()` the argument `values_fill` is a named list specifying what each value should be filled in with when missing. For example:

```{r}
fish_encounters %>%
  pivot_wider(
    names_from = station,
    values_from = seen,
    values_fill = list(seen = 0)
  )
```

In `complete()`, the `fill` argument is a named list that for each variable supplies a single value to use instead of NA for missing combinations. These will also replace existing explicit missing values. 

## Case Study Example

**4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.**

```{r}
who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1) -> who
```

```{r}
who %>%
  group_by(country, year, sex) %>%
  summarise(total = sum(cases)) %>%
  ungroup() %>%
  ggplot(aes(year, total, col = sex)) +
  geom_point()
```



